Biogeography \& environmental conditions shape bacteriophage \& bacteria interaction networks across the human microbiome

### 2017-12-28 doc/oman.tex
Updated manuscript with revisions.

We evaluated the diet-associated differences in gut virome-microbiome
network structure by quantifying how central each sample's network was
on average. We accomplished this by utilizing two common centrality
metrics: degree centrality and closeness centrality. Degree centrality,
the simplest centrality metric, was defined as the number of connections
each phage made with each bacterium. We supplemented measurements of
degree centrality with measurements of closeness centrality. Closeness
centrality is a metric of how close each phage or bacterium is to all of
the other phages and bacteria in the network. A higher closeness
centrality suggests that the effects of genetic information or altered
abundance would be more impactful to all other microbes in the system. A
network with higher average closeness centrality also indicates an
overall greater degree of connections, which suggests a greater
resilience against instability. This is because more highly connected
networks are more stable as a result of pathway dependencies and the
unlikelihood that a randomly removed bacteria or phage would cause major
divisions across the network {[}30,58{]}. We used this information to
calculate the average connectedness per sample, which was corrected for
the maximum potential degree of connectedness. Unfortunately our dataset
was insufficiently powered to make strong conclusions toward this
hypothesis, but this is an interesting observation that warrants further
investigation.

@@

Skin and gut community membership and diversity are highly personal,
with people remaining more similar to themselves than to other people
over time {[}13,59,60{]}. We therefore hypothesized that this personal
conservation extended to microbiome network structure. We addressed this
hypothesis by calculating the degree of dissimilarity between each
subject's network, based on phage and bacteria abundance and centrality.
We quantified phage and bacteria centrality within each sample graph
using the weighted eigenvector centrality metric. This metric defines
central phages as those that are highly abundant (\(A_{O}\) as defined
in the methods) and infect many distinct bacteria which themselves are
abundant and infected by many other phages. Similarly, bacterial
centrality was defined as those bacteria that were both abundant and
connected to numerous phages that were themselves connected to many
bacteria. We then calculated the similarity of community networks using
the weighted eigenvector centrality of all nodes between all samples.
Samples with similar network structures were interpreted as having
similar capacities for maintaining stability and transmitting genetic
material.

@@

Extensive work has illustrated differences in diversity and composition
of the healthy human skin microbiome between anatomical sites, including
bacteria, virus, and fungal communities {[}13,44,59{]}. These
communities vary by degree of skin moisture, oil, and environmental
exposure. As viruses are known to influence microbial diversity and
community composition, we hypothesized that microbe-virus network
structure would be specific to anatomical sites, as well. To test this,
we evaluated the changes in network structure between anatomical sites
within the skin dataset.

@@

Foundational work has provided a baseline understanding of the human
microbiome by characterizing bacterial and viral diversity across the
human body {[}13,14,43--45,61{]}. Here, we offer an initial
understanding of how phage-bacteria networks differ throughout the human
body, so as to provide a baseline for future studies of how and why
microbiome networks differ in disease states. We developed and
implemented a network-based analytical model to evaluate the basic
properties of the human microbiome through bacteria and phage
relationships, instead of membership or diversity alone. This enabled
the application of network theory to provide a new perspective on
complex ecological communities. We utilized metrics of connectivity to
model the extent to which communities of bacteria and phages interact
through mechanisms such as horizontal gene transfer, modulated bacterial
gene expression, and alterations in abundance.

Just as gut microbiome and virome composition and diversity are
conserved in individuals {[}13,43,44,60{]}, gut and skin microbiome
network structures were conserved within individuals over time. Gut
network structure was not conserved among family members. These findings
suggested that the community properties inferred from microbiome
interaction network structures, such as stability (meaning a more highly
connected network is more stable because a randomly removed bacteria or
phage node is less likely to divide or disintegrate {[}30,58{]} the
overall network), the potential for horizontal gene transfer between
members, and co-evolution of populations, were person-specific. These
properties may be impacted by personal factors ranging from the body's
immune system to external environmental conditions, such as climate and
diet.

@@

Second, although our analyses utilized the best datasets currently
available for our study, this work was done retrospectively and relied
on existing data up to seven years old. These archived datasets were
limited by the technology and costs of the time. For example, the diet
and twin studies, relied on multiple displacement amplification (MDA) in
their library preparations--an approach used to overcome the large
nucleic acids requirements typical of older sequencing library
generation protocols. It is now known that MDA results in biases in
microbial community composition {[}62{]}, as well as toward ssDNA viral
genomes {[}63,64{]}, thus rendering the resulting microbial and viral
metagenomes largely non-quantitative. Future work that employs larger
sequence datasets and that avoids the use of bias-inducing amplification
steps will build on and validate our findings, as well as inform the
design and interpretation of further studies.

Finally, the networks in this study were built using operational genomic
units (OGUs), which represented groups of highly similar bacteria or
phage genomes or clustered genome fragments. Similar clustering
definition and validation methods, both computational and experimental,
have been implemented in other metagenomic sequencing studies, as well
{[}42,65--67{]}. These approaches could offer yet another level of
sophistication to our network-based analyses. While this operationally
defined clustering approach allows us to study whole community networks,
our ability to make conclusions about interactions among specific phage
or bacterial species or populations is inherently limited, compared to
more focused, culture-based studies such as the work by Malki \emph{et
al} {[}68{]}. Future work must address this limitation, e.g., through
improved binning methods and deeper metagenomic shotgun sequencing, but
most importantly through an improved conceptual framing of what defines
ecologically and evolutionarily cohesive units for both phage and
bacteria {[}69{]}. Defining operational genomic units and their
taxonomic underpinnings (e.g., whether OGU clusters represent genera or
species) is an active area of work critical to the utility of this
approach. As a first step, phylogenomic analyses have been performed to
cluster cyanophage isolate genomes into informative groups using shared
gene content, average nucleotide identity of shared genes, and pairwise
differences between genomes {[}70{]}. Such population-genetic assessment
of phage evolution, coupled with the ecological implications of genome
heterogeneity, will inform how to define nodes in future iterations of
the ecological network developed here. Even though we are hesitant to
speculate on phage host ranges at low taxonomic levels in our dataset,
the data does aggree with previous reports of instances of broad phage
host range {[}68,71{]}.

Together our work takes an initial step towards defining bacteria-virus
interaction profiles as a characteristic of human-associated microbial
communities. This approach revealed the impacts that different human
environments (e.g., the skin and gut) can have on microbiome
connectivity. By focusing on relationships between bacterial and viral
communities, they are studied as the interacting cohorts they are,
rather than as independent entities. While our developed bacteria-phage
interaction framework is a novel conceptual advance, the microbiome also
consists of archaea and small eukaryotes, including fungi and
\emph{Demodex} mites {[}1,72{]}--all of which can interact with human
immune cells and other non-microbial community members {[}73{]}. Future
work will build from our approach and include these additional community
members and their diverse interactions and relationships (e.g., beyond
phage-bacteria). This will result in a more robust network and a more
holistic understanding of the evolutionary and ecological processes that
drive the assembly and function of the human-associated microbiome.

@@

Raw sequencing data and associated metadata were acquired from the NCBI
sequence read archive (SRA). Supplementary metadata were acquired from
the same SRA repositories and their associated manuscripts. The gut
virome diet study (SRA: \texttt{SRP002424}), twin virome studies (SRA:
\texttt{SRP002523}; \texttt{SRP000319}), and skin virome study (SRA:
\texttt{SRP049645}) were downloaded as \texttt{.sra} files. Sequencing
files were converted to \texttt{fastq} format using the
\texttt{fastq-dump} tool of the NCBI SRA Toolkit (v2.2.0). Sequences
were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases
with quality scores below 33 and shorter than 75 bp {[}74{]}. Paired end
reads were filtered to exclude sequences missing their corresponding
pair using the \texttt{get\_trimmed\_pairs.py} script available in the
source code.

@@

Contigs were assembled using the Megahit assembly program (v1.0.6)
{[}75{]}. A minimum contig length of 1 kb was used. Iterative k-mer
stepping began at a minimum length of 21 and progressed by 20 until 101.
All other default parameters were used.

@@

Contigs were concatenated into two master files prior to alignment, one
for bacterial contigs and one for phage contigs. Sample sequences were
aligned to phage or bacterial contigs using the Bowtie2 global aligner
(v2.2.1) {[}76{]}. We defined a mismatch threshold of 1 bp and seed
length of 25 bp. Sequence abundance was calculated from the Bowtie2
output using the \texttt{calculate\_abundance\_from\_sam.pl} script
available in the source code.

@@

Contigs often represent large fragments of genomes. In order to reduce
redundancy and the resulting artificially inflated genomic richness
within our dataset, it was important to bin contigs into operational
units based on their similarity. This approach is conceptually similar
to the clustering of related 16S rRNA sequences into operational
taxonomic units (OTUs), although here we are clustering contigs into
operational genomic units (OGUs) {[}61{]}.

Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}77{]}.
Because of our large dataset and limits in computational efficiency, we
randomly subsampled the dataset to include 25\% of all samples, and used
these to inform contig abundance within the CONCOCT algorithm. CONCOCT
was used with a maximum of 500 clusters, a k-mer length of four, a
length threshold of 1 kb, 25 iterations, and exclusion of the total
coverage variable.

@@

Open reading frames (ORFs) were identified using the Prodigal program
(V2.6.2) with the meta mode parameter and default settings {[}78{]}.

@@

The four scores were used as parameters in a random forest model to
classify bacteria and bacteriophage pairs as either having infectious
interactions or not. The classification model was built using the Caret
R package (v6.0.73) {[}79{]}. The model was trained using five-fold
cross validation with ten repeats, and the median model performance was
evaluated by training the model on 80\% of the dataset and testing
performance on the remaining 20\%. Pairs without scores were classified
as not interacting. The model was optimized using the ROC value. The
resulting model performance was plotted using the plotROC R package.

@@

Clustered Regularly Interspaced Short Palindromic Repeats (CRISPRs) were
identified from bacterial genomes using the PilerCR program (v1.06)
{[}80{]}. Resulting spacer sequences were filtered to exclude spacers
shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
to the phage genomes using the nucleotide BLAST algorithm with default
parameters (v2.4.0) {[}81{]}. The mean percent identity for each
matching pair was recorded for use in our classification model.

@@

As a result of gene transfer or phage genome integration during
infection, phages may share genes with their bacterial hosts, providing
us with evidence of phage-host pairing. We identified shared genes
between bacterial and phage genomes by assessing amino acid similarity
between the genes using the Diamond protein alignment algorithm
(v0.7.11.60) {[}82{]}. The mean alignment bitscores for each genome pair
were recorded for use in our classification model.

@@

The bacteria and phage operational genomic units (OGUs) were scored
using the same approach as outlined above. The infectious pairings
between bacteria and phage OGUs were classified using the random forest
model described above. The predicted infectious pairings and all
associated metadata were used to populate a graph database using Neo4j
graph database software (v2.3.1) {[}83{]}. This network was used for
downstream community analysis.

@@

Degree and closeness centrality were calculated using the associated
functions within the igraph R package (v1.0.1) {[}84{]}.

@@

\hypertarget{ref-Jeong:2000im}{}
58. Jeong H, Tombor B, Albert R, Oltvai ZN, Barabási AL. The large-scale
organization of metabolic networks. Nature. 2000;407: 651--654.

\hypertarget{ref-Grice:2009ee}{}
59. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
Topographical and Temporal Diversity of the Human Skin Microbiome.
Science. 2009;324: 1190--1192.

\hypertarget{ref-Minot:2013ih}{}
60. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
evolution of the human gut virome. Proceedings of the National Academy
of Sciences of the United States of America. 2013;110: 12450--12455.

\hypertarget{ref-Schloss:2005hz}{}
61. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
defining operational taxonomic units and estimating species richness.
Applied and Environmental Microbiology. 2005;71: 1501--1506.

\hypertarget{ref-Yilmaz:2010jb}{}
62. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
amplification compromises quantitative analysis of metagenomes. Nature
Methods. 2010;7: 943--944.

\hypertarget{ref-Kim:2008to}{}
63. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
single-stranded DNA viruses from rice paddy soil. Applied and \ldots{}.
2008;

\hypertarget{ref-Kim:2011hp}{}
64. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
of uncultured single-stranded and double-stranded DNA viruses. Applied
and Environmental Microbiology. 2011;77: 7663--7668.

\hypertarget{ref-Minot:2012ed}{}
65. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
among diverse viruses of the human gut. PLOS ONE. 2012;7: e42342.

\hypertarget{ref-Deng:2014eb}{}
66. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
Hugenholtz P, et al. Viral tagging reveals discrete populations in
Synechococcus viral genome sequence space. Nature. 2014;513: 242--245.

\hypertarget{ref-Brum:2015iaa}{}
67. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
A, et al. Ocean plankton. Patterns and ecological drivers of ocean viral
communities. Science. 2015;348: 1261498--1261498.

\hypertarget{ref-Malki:2015dw}{}
68. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
Bacteriophages isolated from Lake Michigan demonstrate broad host-range
across several bacterial phyla. Virology Journal. 2015;12: 164.

\hypertarget{ref-Polz:2006fi}{}
69. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
of genetic and phenotypic differentiation in marine microbes.
Philosophical Transactions of the Royal Society B: Biological Sciences.
2006;361: 2009--2021.

\hypertarget{ref-Gregory:2016cg}{}
70. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
A, Sudek S, et al. Genomic differentiation among wild cyanophages
despite widespread horizontal gene transfer. BMC Genomics. 2016;17: 930.

\hypertarget{ref-PaezEspino:2016gi}{}
71. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
M, Mikhailova N, et al. Uncovering Earth's virome. Nature. 2016;

\hypertarget{ref-Grice:2011gy}{}
72. Grice EA, Segre JA. The skin microbiome. Nature Reviews
Microbiology. 2011;9: 244--253.

\hypertarget{ref-Round:2009bz}{}
73. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
responses during health and disease. Nature reviews Immunology. 2009;9:
313--323.

\hypertarget{ref-FASTXToolkit:wr}{}
74. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.

\hypertarget{ref-Li:2016kd}{}
75. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
MEGAHIT v1.0: A fast and scalable metagenome assembler driven by
advanced methodologies and community practices. METHODS. 2016;102:
3--11.

\hypertarget{ref-Langmead:2012jh}{}
76. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
Nature Methods. 2012;9: 357--359.

\hypertarget{ref-Alneberg:2014fc}{}
77. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
J, Ijaz UZ, et al. Binning metagenomic contigs by coverage and
composition. Nature Methods. 2014; 1--7.

\hypertarget{ref-Hyatt:2012cy}{}
78. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
initiation site prediction in metagenomic sequences. Bioinformatics.
2012;28: 2223--2230.

\hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
79. Kuhn M. caret: Classification and Regression Training.

\hypertarget{ref-Edgar:2007bh}{}
80. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
repeats. BMC Bioinformatics. 2007;8: 18.

\hypertarget{ref-Camacho:2009fc}{}
81. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
et al. BLAST+: architecture and applications. BMC Bioinformatics.
2009;10: 1.

\hypertarget{ref-Buchfink:2015ki}{}
82. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
using DIAMOND. Nature Methods. 2015;12: 59--60.

\hypertarget{ref-Neoj:Fuwr6PBN}{}
83. Neo4j.

\hypertarget{ref-Theigraphsoftware:vh}{}
84. Csardi G, Nepusz T. The igraph software package for complex network
research.

>>>>

We evaluated the diet-associated differences in gut virome-microbiome
network structure by quantifying how central each sample's network was
on average. We accomplished this by utilizing two common centrality
metrics: degree centrality and closeness centrality. Degree centrality,
the simplest centrality metric, was defined as the number of connections
each phage made with each bacterium. We supplemented measurements of
degree centrality with measurements of closeness centrality. Closeness
centrality is a metric of how close each phage or bacterium is to all of
the other phages and bacteria in the network. A higher closeness
centrality suggests that the effects of genetic information or altered
abundance would be more impactful to all other microbes in the system. A
network with higher average closeness centrality also indicates an
overall greater degree of connections, which suggests a greater
resilience against instability. This is because more highly connected
networks are more stable as a result of pathway dependencies and the
unlikelihood that a randomly removed bacteria or phage would cause major
divisions across the network {[}{\textbf{???}},30{]}. We used this
information to calculate the average connectedness per sample, which was
corrected for the maximum potential degree of connectedness.
Unfortunately our dataset was insufficiently powered to make strong
conclusions toward this hypothesis, but this is an interesting
observation that warrants further investigation.

@@

Skin and gut community membership and diversity are highly personal,
with people remaining more similar to themselves than to other people
over time {[}13,58,59{]}. We therefore hypothesized that this personal
conservation extended to microbiome network structure. We addressed this
hypothesis by calculating the degree of dissimilarity between each
subject's network, based on phage and bacteria abundance and centrality.
We quantified phage and bacteria centrality within each sample graph
using the weighted eigenvector centrality metric. This metric defines
central phages as those that are highly abundant (\(A_{O}\) as defined
in the methods) and infect many distinct bacteria which themselves are
abundant and infected by many other phages. Similarly, bacterial
centrality was defined as those bacteria that were both abundant and
connected to numerous phages that were themselves connected to many
bacteria. We then calculated the similarity of community networks using
the weighted eigenvector centrality of all nodes between all samples.
Samples with similar network structures were interpreted as having
similar capacities for maintaining stability and transmitting genetic
material.

@@

Extensive work has illustrated differences in diversity and composition
of the healthy human skin microbiome between anatomical sites, including
bacteria, virus, and fungal communities {[}13,44,58{]}. These
communities vary by degree of skin moisture, oil, and environmental
exposure. As viruses are known to influence microbial diversity and
community composition, we hypothesized that microbe-virus network
structure would be specific to anatomical sites, as well. To test this,
we evaluated the changes in network structure between anatomical sites
within the skin dataset.

@@

Foundational work has provided a baseline understanding of the human
microbiome by characterizing bacterial and viral diversity across the
human body {[}13,14,43--45,60{]}. Here, we offer an initial
understanding of how phage-bacteria networks differ throughout the human
body, so as to provide a baseline for future studies of how and why
microbiome networks differ in disease states. We developed and
implemented a network-based analytical model to evaluate the basic
properties of the human microbiome through bacteria and phage
relationships, instead of membership or diversity alone. This enabled
the application of network theory to provide a new perspective on
complex ecological communities. We utilized metrics of connectivity to
model the extent to which communities of bacteria and phages interact
through mechanisms such as horizontal gene transfer, modulated bacterial
gene expression, and alterations in abundance.

Just as gut microbiome and virome composition and diversity are
conserved in individuals {[}13,43,44,59{]}, gut and skin microbiome
network structures were conserved within individuals over time. Gut
network structure was not conserved among family members. These findings
suggested that the community properties inferred from microbiome
interaction network structures, such as stability (meaning a more highly
connected network is more stable because a randomly removed bacteria or
phage node is less likely to divide or disintegrate
{[}{\textbf{???}},30{]} the overall network), the potential for
horizontal gene transfer between members, and co-evolution of
populations, were person-specific. These properties may be impacted by
personal factors ranging from the body's immune system to external
environmental conditions, such as climate and diet.

@@

Second, although our analyses utilized the best datasets currently
available for our study, this work was done retrospectively and relied
on existing data up to seven years old. These archived datasets were
limited by the technology and costs of the time. For example, the diet
and twin studies, relied on multiple displacement amplification (MDA) in
their library preparations--an approach used to overcome the large
nucleic acids requirements typical of older sequencing library
generation protocols. It is now known that MDA results in biases in
microbial community composition {[}61{]}, as well as toward ssDNA viral
genomes {[}62,63{]}, thus rendering the resulting microbial and viral
metagenomes largely non-quantitative. Future work that employs larger
sequence datasets and that avoids the use of bias-inducing amplification
steps will build on and validate our findings, as well as inform the
design and interpretation of further studies.

Finally, the networks in this study were built using operational genomic
units (OGUs), which represented groups of highly similar bacteria or
phage genomes or clustered genome fragments. Similar clustering
definition and validation methods, both computational and experimental,
have been implemented in other metagenomic sequencing studies, as well
{[}42,64--66{]}. These approaches could offer yet another level of
sophistication to our network-based analyses. While this operationally
defined clustering approach allows us to study whole community networks,
our ability to make conclusions about interactions among specific phage
or bacterial species or populations is inherently limited, compared to
more focused, culture-based studies such as the work by Malki \emph{et
al} {[}67{]}. Future work must address this limitation, e.g., through
improved binning methods and deeper metagenomic shotgun sequencing, but
most importantly through an improved conceptual framing of what defines
ecologically and evolutionarily cohesive units for both phage and
bacteria {[}68{]}. Defining operational genomic units and their
taxonomic underpinnings (e.g., whether OGU clusters represent genera or
species) is an active area of work critical to the utility of this
approach. As a first step, phylogenomic analyses have been performed to
cluster cyanophage isolate genomes into informative groups using shared
gene content, average nucleotide identity of shared genes, and pairwise
differences between genomes {[}69{]}. Such population-genetic assessment
of phage evolution, coupled with the ecological implications of genome
heterogeneity, will inform how to define nodes in future iterations of
the ecological network developed here. Even though we are hesitant to
speculate on phage host ranges at low taxonomic levels in our dataset,
the data does aggree with previous reports of instances of broad phage
host range {[}67,70{]}.

Together our work takes an initial step towards defining bacteria-virus
interaction profiles as a characteristic of human-associated microbial
communities. This approach revealed the impacts that different human
environments (e.g., the skin and gut) can have on microbiome
connectivity. By focusing on relationships between bacterial and viral
communities, they are studied as the interacting cohorts they are,
rather than as independent entities. While our developed bacteria-phage
interaction framework is a novel conceptual advance, the microbiome also
consists of archaea and small eukaryotes, including fungi and
\emph{Demodex} mites {[}1,71{]}--all of which can interact with human
immune cells and other non-microbial community members {[}72{]}. Future
work will build from our approach and include these additional community
members and their diverse interactions and relationships (e.g., beyond
phage-bacteria). This will result in a more robust network and a more
holistic understanding of the evolutionary and ecological processes that
drive the assembly and function of the human-associated microbiome.

@@

Raw sequencing data and associated metadata were acquired from the NCBI
sequence read archive (SRA). Supplementary metadata were acquired from
the same SRA repositories and their associated manuscripts. The gut
virome diet study (SRA: \texttt{SRP002424}), twin virome studies (SRA:
\texttt{SRP002523}; \texttt{SRP000319}), and skin virome study (SRA:
\texttt{SRP049645}) were downloaded as \texttt{.sra} files. Sequencing
files were converted to \texttt{fastq} format using the
\texttt{fastq-dump} tool of the NCBI SRA Toolkit (v2.2.0). Sequences
were quality trimmed using the Fastx toolkit (v0.0.14) to exclude bases
with quality scores below 33 and shorter than 75 bp {[}73{]}. Paired end
reads were filtered to exclude sequences missing their corresponding
pair using the \texttt{get\_trimmed\_pairs.py} script available in the
source code.

@@

Contigs were assembled using the Megahit assembly program (v1.0.6)
{[}74{]}. A minimum contig length of 1 kb was used. Iterative k-mer
stepping began at a minimum length of 21 and progressed by 20 until 101.
All other default parameters were used.

@@

Contigs were concatenated into two master files prior to alignment, one
for bacterial contigs and one for phage contigs. Sample sequences were
aligned to phage or bacterial contigs using the Bowtie2 global aligner
(v2.2.1) {[}75{]}. We defined a mismatch threshold of 1 bp and seed
length of 25 bp. Sequence abundance was calculated from the Bowtie2
output using the \texttt{calculate\_abundance\_from\_sam.pl} script
available in the source code.

@@

Contigs often represent large fragments of genomes. In order to reduce
redundancy and the resulting artificially inflated genomic richness
within our dataset, it was important to bin contigs into operational
units based on their similarity. This approach is conceptually similar
to the clustering of related 16S rRNA sequences into operational
taxonomic units (OTUs), although here we are clustering contigs into
operational genomic units (OGUs) {[}60{]}.

Contigs were clustered using the CONCOCT algorithm (v0.4.0) {[}76{]}.
Because of our large dataset and limits in computational efficiency, we
randomly subsampled the dataset to include 25\% of all samples, and used
these to inform contig abundance within the CONCOCT algorithm. CONCOCT
was used with a maximum of 500 clusters, a k-mer length of four, a
length threshold of 1 kb, 25 iterations, and exclusion of the total
coverage variable.

@@

Open reading frames (ORFs) were identified using the Prodigal program
(V2.6.2) with the meta mode parameter and default settings {[}77{]}.

@@

The four scores were used as parameters in a random forest model to
classify bacteria and bacteriophage pairs as either having infectious
interactions or not. The classification model was built using the Caret
R package (v6.0.73) {[}78{]}. The model was trained using five-fold
cross validation with ten repeats, and the median model performance was
evaluated by training the model on 80\% of the dataset and testing
performance on the remaining 20\%. Pairs without scores were classified
as not interacting. The model was optimized using the ROC value. The
resulting model performance was plotted using the plotROC R package.

@@

Clustered Regularly Interspaced Short Palindromic Repeats (CRISPRs) were
identified from bacterial genomes using the PilerCR program (v1.06)
{[}79{]}. Resulting spacer sequences were filtered to exclude spacers
shorter than 20 bp and longer than 65 bp. Spacer sequences were aligned
to the phage genomes using the nucleotide BLAST algorithm with default
parameters (v2.4.0) {[}80{]}. The mean percent identity for each
matching pair was recorded for use in our classification model.

@@

As a result of gene transfer or phage genome integration during
infection, phages may share genes with their bacterial hosts, providing
us with evidence of phage-host pairing. We identified shared genes
between bacterial and phage genomes by assessing amino acid similarity
between the genes using the Diamond protein alignment algorithm
(v0.7.11.60) {[}81{]}. The mean alignment bitscores for each genome pair
were recorded for use in our classification model.

@@

The bacteria and phage operational genomic units (OGUs) were scored
using the same approach as outlined above. The infectious pairings
between bacteria and phage OGUs were classified using the random forest
model described above. The predicted infectious pairings and all
associated metadata were used to populate a graph database using Neo4j
graph database software (v2.3.1) {[}82{]}. This network was used for
downstream community analysis.

@@

Degree and closeness centrality were calculated using the associated
functions within the igraph R package (v1.0.1) {[}83{]}.

@@

\hypertarget{ref-Grice:2009ee}{}
58. Grice EA, Kong HH, Conlan S, Deming CB, Davis J, Young AC, et al.
Topographical and Temporal Diversity of the Human Skin Microbiome.
Science. 2009;324: 1190--1192.

\hypertarget{ref-Minot:2013ih}{}
59. Minot S, Bryson A, Chehoud C, Wu GD, Lewis JD, Bushman FD. Rapid
evolution of the human gut virome. Proceedings of the National Academy
of Sciences of the United States of America. 2013;110: 12450--12455.

\hypertarget{ref-Schloss:2005hz}{}
60. Schloss PD, Handelsman J. Introducing DOTUR, a computer program for
defining operational taxonomic units and estimating species richness.
Applied and Environmental Microbiology. 2005;71: 1501--1506.

\hypertarget{ref-Yilmaz:2010jb}{}
61. Yilmaz S, Allgaier M, Hugenholtz P. Multiple displacement
amplification compromises quantitative analysis of metagenomes. Nature
Methods. 2010;7: 943--944.

\hypertarget{ref-Kim:2008to}{}
62. Kim KH, Chang HW, Nam YD, Roh SW. Amplification of uncultured
single-stranded DNA viruses from rice paddy soil. Applied and \ldots{}.
2008;

\hypertarget{ref-Kim:2011hp}{}
63. Kim K-H, Bae J-W. Amplification methods bias metagenomic libraries
of uncultured single-stranded and double-stranded DNA viruses. Applied
and Environmental Microbiology. 2011;77: 7663--7668.

\hypertarget{ref-Minot:2012ed}{}
64. Minot S, Wu GD, Lewis JD, Bushman FD. Conservation of gene cassettes
among diverse viruses of the human gut. PLOS ONE. 2012;7: e42342.

\hypertarget{ref-Deng:2014eb}{}
65. Deng L, Ignacio-Espinoza JC, Gregory AC, Poulos BT, Weitz JS,
Hugenholtz P, et al. Viral tagging reveals discrete populations in
Synechococcus viral genome sequence space. Nature. 2014;513: 242--245.

\hypertarget{ref-Brum:2015iaa}{}
66. Brum JR, Ignacio-Espinoza JC, Roux S, Doulcier G, Acinas SG, Alberti
A, et al. Ocean plankton. Patterns and ecological drivers of ocean viral
communities. Science. 2015;348: 1261498--1261498.

\hypertarget{ref-Malki:2015dw}{}
67. Malki K, Kula A, Bruder K, Sible E, Hatzopoulos T, Steidel S, et al.
Bacteriophages isolated from Lake Michigan demonstrate broad host-range
across several bacterial phyla. Virology Journal. 2015;12: 164.

\hypertarget{ref-Polz:2006fi}{}
68. Polz MF, Hunt DE, Preheim SP, Weinreich DM. Patterns and mechanisms
of genetic and phenotypic differentiation in marine microbes.
Philosophical Transactions of the Royal Society B: Biological Sciences.
2006;361: 2009--2021.

\hypertarget{ref-Gregory:2016cg}{}
69. Gregory AC, Solonenko SA, Ignacio-Espinoza JC, LaButti K, Copeland
A, Sudek S, et al. Genomic differentiation among wild cyanophages
despite widespread horizontal gene transfer. BMC Genomics. 2016;17: 930.

\hypertarget{ref-PaezEspino:2016gi}{}
70. Paez-Espino D, Eloe-Fadrosh EA, Pavlopoulos GA, Thomas AD, Huntemann
M, Mikhailova N, et al. Uncovering Earth's virome. Nature. 2016;

\hypertarget{ref-Grice:2011gy}{}
71. Grice EA, Segre JA. The skin microbiome. Nature Reviews
Microbiology. 2011;9: 244--253.

\hypertarget{ref-Round:2009bz}{}
72. Round JL, Mazmanian SK. The gut microbiota shapes intestinal immune
responses during health and disease. Nature reviews Immunology. 2009;9:
313--323.

\hypertarget{ref-FASTXToolkit:wr}{}
73. Hannon GJ. FASTX-Toolkit.: GNU Affero General Public License.

\hypertarget{ref-Li:2016kd}{}
74. Li D, Luo R, Liu C-M, Leung C-M, Ting H-F, Sadakane K, et al.
MEGAHIT v1.0: A fast and scalable metagenome assembler driven by
advanced methodologies and community practices. METHODS. 2016;102:
3--11.

\hypertarget{ref-Langmead:2012jh}{}
75. Langmead B, Salzberg SL. Fast gapped-read alignment with Bowtie 2.
Nature Methods. 2012;9: 357--359.

\hypertarget{ref-Alneberg:2014fc}{}
76. Alneberg J, Bjarnason BS aacute ri, Bruijn I de, Schirmer M, Quick
J, Ijaz UZ, et al. Binning metagenomic contigs by coverage and
composition. Nature Methods. 2014; 1--7.

\hypertarget{ref-Hyatt:2012cy}{}
77. Hyatt D, LoCascio PF, Hauser LJ, Uberbacher EC. Gene and translation
initiation site prediction in metagenomic sequences. Bioinformatics.
2012;28: 2223--2230.

\hypertarget{ref-caretClassificatio:ux5fU2Litux5f1}{}
78. Kuhn M. caret: Classification and Regression Training.

\hypertarget{ref-Edgar:2007bh}{}
79. Edgar RC. PILER-CR: fast and accurate identification of CRISPR
repeats. BMC Bioinformatics. 2007;8: 18.

\hypertarget{ref-Camacho:2009fc}{}
80. Camacho C, Coulouris G, Avagyan V, Ma N, Papadopoulos J, Bealer K,
et al. BLAST+: architecture and applications. BMC Bioinformatics.
2009;10: 1.

\hypertarget{ref-Buchfink:2015ki}{}
81. Buchfink B, Xie C, Huson DH. Fast and sensitive protein alignment
using DIAMOND. Nature Methods. 2015;12: 59--60.

\hypertarget{ref-Neoj:Fuwr6PBN}{}
82. Neo4j.

\hypertarget{ref-Theigraphsoftware:vh}{}
83. Csardi G, Nepusz T. The igraph software package for complex network
research.
